Mathematical Aspects of the Capital Market

### 2017-07-24 phd/Chapterintroduction.tex
Revised the introduction section of the thesis. Started programming for the main importance sampling algorithm.

One of the most important dynamical mechanisms that lead to regularly
varying r.v. is stochastic recursion of the following form:
\[
X_t = A_t X_{t-1} + B_t
\]
where $X_t$ is a $d$-dimensional random vector, $A_t$ is a $d\times d$
random matrix and $B_t$ is a $d$-dimensional vector, random or
deterministic.

\section{Stochastic Recurrence Equation}

>>>>

\section{Stochastic Recurrence Equation}
One of the most important dynamical mechanisms that lead to regularly
varying r.v. is stochastic recursion of the following form:
\[
X_t = A_t X_{t-1} + B_t
\]
where $X_t$ is a $d$-dimensional random vector, $A_t$ is a $d\times d$
random matrix and $B_t$ is a $d$-dimensional vector, random or
deterministic. The sequence $\{A_i\}_{i=1,2,\dots}$ and
$\{B_i\}_{i=1,2,\dots}$  are iid and independent of each other.

Kesten \cite{kesten:1973} showed that, when $A_t$ and $B_t$ were almost
surely non-negative, had no row or column of only zeros, and there was
a positive probability that $B_t$ was strictly positive, the strictly
stationary solution to the equation
$V \overset{d}{=} A V + B$ had power-law tails
for its marginal distributions, assuming the following conditions (M)
and (A):
\begin{itemize}
\item Condition (M)
  \begin{enumerate}
  \item The top Lyapunov exponent
    \[
    \gamma = \inf_{n \geq 1} {1 \over n}\E \log \|A_n \cdots A_1\|
    \]
    is negative.
  \item There exists $\xi > 0$ such that
    \[
    1 = \lambda(\xi) = \lim_{n \to \infty} {1 \over n} \log \E \|A_n \cdots A_1\|^\xi
    \]
  \item $\E (\|A_1\|^\xi \log^+\|A_1\|) < \infty$
  \item $\E |B_1|^\xi < \infty$
  \end{enumerate}
\item Condition (A) : The group generated by
  \[
  \{\log\rho(s): s = A_n \cdots A_1 \text{ for some } n \geq 1\}
  \]
  is dense in $\reals$, where $\rho(s)$ denotes the spectral
  radius of matrix $s$.
\end{itemize}
Upon these conditions, Kesten's theorem gives
\begin{equation*}
  u^\xi \P(u^{-1} V \in \cdot) \overset{v}{\to} \mu_\xi(\cdot)
\end{equation*}
where $\mu_\xi$ is a non-null Radon measure on
$\reals^d_+ \setminus \{0\}$ with the property
$\mu_\xi(a A) = a^{-\xi} \mu_\xi(A)$.

Recently, Collamore and Mentemeier \cite{collamore:mentemeier:2016}
extended Kesten's result and gave an explicit expression for $\mu_\xi$:
\begin{equation}
  \label{eq:CollamoreMentemeierIntro}
  \lim_{u \to \infty} u^{\xi} \E \left[
    f(u^{-1} V)
    \right]
  =
  {C \over \lambda'(\xi)}  
  \int_{\sphere^{d-1}_+ \times R} e^{-\xi s} f(e^s x) \ell_\xi(dx) ds
\end{equation}
where $C$ is a constant (cf. eq.(2.15) of Collamore and Mentemeier
\cite{collamore:mentemeier:2016}), $f(\cdot)$ is any bounded
continuous function on $\reals^{d}_+ \setminus \{0\}$  and
$\ell_\xi$ is a probability measure on $\sphere_+^{d-1}$. Its definition
is also found in \eqref{eq:eigenmeasure}.

From \eqref{eq:CollamoreMentemeierIntro} a representation for $\mu_\xi
(\cdot)$ immediately follows  
\[
\mu_\xi (\cdot) = {C \over \lambda'(\xi)} \mathcal L_\xi(\cdot)
\]
Here $\mathcal L_\xi$ is a non-null Radon measure  on
$\reals^d_+ \setminus \{0\}$ that satisfies, for all
bounded continuous function $f(\cdot)$ on
$\reals_+^{d} \setminus \{0\}$:
\[
\int_{\reals_+^d\setminus \{0\}} f(x) \mathcal L_\xi(dx)
=
\int_{\sphere^{d-1}_+ \times R} e^{-\xi s} f(e^s x) \ell_\xi(dx) ds
\]

### 2017-07-24 phd/MyThesis.tex
Revised the introduction section of the thesis. Started programming for the main importance sampling algorithm.

\input{Preamble}
\usepackage[numbers,sort&compress]{natbib}

@@

\begin{tabular}{ll}
Supervisor: & Prof.~Thomas Mikosch\\
& University of Copenhagen\\
& \\
Co-supervisor: & Associate Prof.~Jeffrey Collamore\\
& University of Copenhagen\\
& \\
Assessment committee: &  Prof.~Catalin Starica\\
& University of Neuchatel, CH\\
& \\

@@

\def\phdthesis{1}
\input{Chapterintroduction}
%------------------------------------------------------------------------------------
\input{ChapterTailParameters}
%% \input{TailParameters}
\input{Chapterextremes}

>>>>

\input{Preamble}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{algorithm}
\usepackage{algpseudocode}

\newtheorem{case}{case}

@@

\begin{tabular}{ll}
Supervisor: & Prof.~Thomas Mikosch\\
& University of Copenhagen\\
& \\
Co-supervisor: & Associate Prof.~Jeffrey Collamore\\
& University of Copenhagen\\
& \\
Assessment committee: &  Prof.~Catalin St{\u{a}}ric{\u{a}}\\
& University of Neuchatel, CH\\
& \\

@@

\def\phdthesis{1}
\input{Chapterintroduction}
%------------------------------------------------------------------------------------
\input{ChapterTailParameters}
\input{ChapterImportanceSampling}
\input{Chapterextremes}
